---
title: "R-Programming Lab \t : Assignment 3"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```




\newpage
***
**Question1**

Find the minimum of the function  f(x)=x1-x2+2*x1^2+2*x1*x2+x2^2
using Newton's method.



**Libraries**

```{r message=FALSE, warning=FALSE}
library(numDeriv)
library(pracma)
```


**Defining function**

**f(x)**

```{r}

func<- function(x) {
  x1<-x[1]
  x2<-x[2]
  
  result<-x1-x2+2*x1^2+2*x1*x2+x2^2
  result
}


```



**Hessian Matrix of f(x)**


```{r}

hessian_mat=hessian(func,c(0,0))
hessian_matrix<-as.matrix(hessian_mat)
cat("Hessian of function f(x) is:")
hessian_matrix


```

 **Inverse of hessian matrix**

```{r}
H=inv(hessian_matrix)   #Inverse of hessian matrix
H
```

**tolerance and iteration**

```{r}
tolerance <- c(0,0)  #set tolerance=0 (when gradient of function is zero  Then x is optimum point)
k=0  #Here k is iteration
```



\newpage
***

**code of newton method for finding minima of function**
```{r}
root<-function(x,g,tolerance){
  while( (all(abs(g)>=tolerance)) ){
    if(k <101){
      g=x
      x=x-H %*% grad(func,g)
      k=k+1
      #print(k)
      #print("\n")
      #print(x)
      
    }
    else{
      break;
    }
    
    
    
  }
  return(x)
}
```



**choosing randomly 100 guess**

```{r}
a=seq(0,4.95,0.05)
b=seq(0,4.95,0.05)
```

```{r}
data.frame(a,b)
```

Each row in dataframe represents initial value of x.
And from above dataframe you can see that there are  100 differents points 
and point (0,0) is included at index 1

\newpage
***
```{r}

m<-matrix(,nrow=2,ncol=100)


```

**Finding minima of function for 100 points**  

```{r}


for(i in 1:100){
  x<-c(a[i],b[i])
  m[,i]=root(x,grad(func,x),tolerance)
  
}
```

**minimum value of f(x)**
```{r}
m
```



from above matrix(2*100) you can observe that the 
Each column represents the minimum of  function calculating using newton's method (implemented from scratch) for different starting values .  
we are taken 100 such differnet values and minimum came out to be the same value for each case .



\newpage
***
**Value of function f(x) at mimima (x1=-1,x2=1.5)**

```{r}
ans<-0
for(i in 1:100){
   ans[i]<-func(c(m[,i]))
   
}
```

```{r}
ans
```


\newpage
***
**using inbuilt function(nleqslv)**

**Finding minima function using library function**






```{r}

f <- function(x){
  y <- numeric(2)
  y[1] <- 4*x[1] + 2*x[2]+1
  y[2] <-  2*x[1] + 2*x[2]- 1
  y
}
```




```{r}
jacob <- function(x) {
  n <- length(x)
  Df <- matrix(numeric(n*n),n,n)
  Df[1,1] <- 4
  Df[1,2] <- 2
  Df[2,1] <- 2
  Df[2,2] <- 2
  Df
}
```



```{r}
a=seq(-1,4.95,0.06)
b=seq(-1,4.95,0.06)
library(numDeriv)
library(pracma)

library(nleqslv)
mat<-matrix(,nrow=2,ncol=100)


for(i in 1:100){
  xstart <- c(a[i],b[i]) 
  fstart <- jacob(xstart)
  xstart
  fstart
  mat[,i]<-nleqslv(xstart,f,method="Newton")$x
  
}

```
\newpage
***
**minimum value of f(x) **


```{r}
mat
```


from above matrix(2*100) you can observe that the 
Each column represents the minimum of  function calculating using newton's method (implemented using inbuilt function) for different starting values .  
we are taken 100 such differnet values and minimum came out to be the same value for each case .

\newpage
***

**Value of function f(x) at minima (using inbuilt function)**


```{r}
ans_inbuilt<-0
for(i in 1:100){
   ans_inbuilt[i]<-func(c(mat[,i]))
   
}
   
```


```{r}
ans_inbuilt
```
\newpage
***
**Solution Agreement with library function**
```{r}
own_ans<-ans
lib_ans<-ans_inbuilt
```



```{r}
data.frame(own_ans,lib_ans)
```

from above dataframe you can observe that the data in column 1(own_ans) is exactly equal to column 2(lib_ans) as expected.




